Different forms of uncertainty differently affect the evolution of social learning

### 2022-07-15 ms.tex
Updates from Overleaf

\usepackage{graphicx}

@@

\section{Introduction}
\cm{Notes on our June 2 discussion (USPs): "Uncertainty favors social learning" is a widespread belief. But it is operationalized in different ways, and oftentimes it does not reflect REAL uncertainty. Selling points of this paper 1) real uncertainty, 2) 4 kinds uncertainty, 3) cognition (soft-max temperature varies how much explores).}

%% for inclusion in the social learning ms

Decision-making under uncertainty is fundamental challenge for any organism. Uncertainty weighs particularly heavily on humans adaptation because of our long lifespans, cosmopolitan distribution, and the resulting necessity to adapt to relatively coarse-grain environments \cite{levins1962}.

While the importance of decision-making under uncertainty is generally acknowledged, there is often some confusion about what constitutes uncertainty. First, uncertainty needs to be distinguished from risk. While the concepts sound similar, uncertainty and risk are quite different. Both refer to the existence of non-deterministic outcomes, but with risk, outcomes can be represented probabilistically, whereas uncertainty implies that the probabilities of outcomes remain unknown \cite{knight1921}. As it doesn't lend itself to simple mathematical representation, uncertainty presents substantial analytical challenges as first noted by \cite{keynes1921}.

Despite the technical difficulties of dealing with uncertainty, addressing it is an essential task because it potentially changes our understanding of decision-making in a qualitative manner. The dominant approach to understanding decisions for choices with variable outcomes is expected utility theory (EUT). As the name suggests, EUT requires the ability to calculate expectations, which means that we must know the probability distributions of outcomes. The non-probabilistic nature of uncertainty means one cannot calculate expectations. Following \cite{savage1954}, various research traditions in decision theory and economics have treated uncertainty as if it were risk, just with subjective probabilities replacing objective ones. The fundamental problem with this approach was noted by \cite{geweke2001} and further developed by \cite{weitzman2009}. In particular, the probability density function of an outcome that needs to be learned, but for which opportunities for learning are highly constrained, will be heavy-tailed and expected-utility calculations will not work because of the resulting undefined mass-generating function. Uncertainty implies heavy-tailed probability distributions and heavy-tailed distributions require different analytical methods \cite{nair_etal2022}.

The heuristic-based adaptive-toolbox approach of \cite{todd_gigerenzer2000} provides an alternative strategy for understanding decision-making under uncertainty. There is a strong intuition that uncertainty should activate specific social-learning heuristics. Within the cultural-evolution literature, uncertainty is generally thought to lead to a social heuristic known as \emph{conformist-biased learning} \cite{boyd_richerson1985, henrich_boyd1998, muthukrishna_etal2016}.

True uncertainty arises when an organism lacks the capacity to learn about the possible outcomes of a decision task. This is very likely for a cosmopolitan species living in a coarse-grain environment. Task complexity also. Non-stationary environmental change \cite{kay_king2020}. These potential sources of uncertainty can also interact.

\cm{CUT in favor of uncertainty intro: Social learning enhances problem solving when acquiring information from others is more efficient than learning on one's own~\cite{Laland2004}. However, social learning can also lead individuals astray if, for example, they are copying outdated information. Theoretical models have helped clarify the circumstances under which social learning should be evolutionarily favored. For example, environments that change across time favor learning from others (as opposed to having a genetically fixed behavior) ~\cite{BoydRicherson1985,Henrich1998}, so long as the environments are not so variable that the social information reflects outdated and maladaptive behavior point~\cite{Rogers1988,Feldman1996, aoki2014evolution}. \cm{check that the citations still accurate}. This theoretical work has motivated much empirical research and helped predict social learning behavior in humans and other species ~\cite{McElreath2005,Kendal2018,Allen2019}.
}

From this literature, the idea that uncertainty supports social learning has proliferated, but the term ``uncertainty" often conflates several different interpretations of that word. For example, different models define uncertainty in terms of the rate of environmental change, the correlation between spatial environments, or the reliability of information acquired from the environment. 
Furthermore, in translating mathematical models to verbal predictions, the formal meaning of terms like uncertainty is often lost \cite{lawson1988probability}, facilitating such conflation. 
%Furthermore, a
And as the number of formal models of social learning have expanded, an ever increasing number
of modeling choices ~\cite[Figure 1]{Kendal2018} and formalizations of uncertainty have made it difficult to compare across models and consolidate our understanding of the contexts in which social learning is adaptive.

To address these concerns we developed an empirically-motivated, evolutionary agent-based model that simultaneously operationalizes four %principal components 
different forms of uncertainty: (1) temporal environmental variability; (2) selection set size (the number of possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for different behaviors); and (4) the effective lifespan of agents (the number of opportunities for individual learning). These were chosen because modeling and empirical 
studies across taxa tend use one or more of these uncertainty dimensions\ps{citation needed}.  
The model also involves a more cognitively realistic learning mechanism, the softmax algorithm, which allows us to more accurately assess some of the likely tradeoffs between social and individual learning. We use this integrated model to ask what kinds of uncertainty are likely to favor social learning, and to clarify the logic of why this would be the case. Answering this question is critical for understanding human behavior adapting to a rapidly changing world, both to mitigate existential threats~\cite{Moya2020,Jones2021}, and to capitalize on new opportunities such as transitioning to clean energy use and production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.

In the remainder of this introduction, we further unpack our motivation for this work justify our approach. We then present our model and the results of our analyses. We conclude with a discussion of our results and their implications for our understanding of social learning, uncertainty, and cultural evolution.

@@

\textbf{Temporal environmental variability.}
When external or internal shocks (e.g., climatic events, migration, technological paradigm shifts) occur, strategies that were previously adaptive may no longer be optimal. The difficulty in predicting either when such shocks will occur or what behaviors will adaptive in the resulting environments leads to uncertainty. 
Temporal environmental variability is fundamental to most evolutionary models of social learning, yet even this is implemented in several ways. It is usually modeled as an independent probability of the environment changing its state each generation ~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}. The probabilities of environmental change are usually fixed, %~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
but they have also been modeled as deterministic cycles ~\cite{Feldman1996, aoki2014evolution} or as variable rates \cm{who does this?}\ps{Yeah, I'm not even sure what variable rates means in this context}. Furthermore, the consequences of environmental change can range from mild to  catastrophic. In the latter case, a change of environment results in the total elimination of any adaptive behavior, which must be learned \emph{de novo} for the new environment ~\cite{Rogers1988}. Such catastrophic environmental changes present a large adaptive challenge as individuals cannot bank on accumulated information from previous generations. Other models of environmental change introduce less uncertainty as they fluctuate between two environmental states with corresponding adaptive behaviors. This means that all individuals with previously maladaptive behaviors then have adaptive behaviors upon the environment changing ~\cite{perreault2012bayesian}.

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options. This is a mathematical fact at the core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the selection set doubles in size  \cite{mackay2003information}. 
This is intuitive; the more options one could potentially investigate, the greater one's uncertainty about which option is the best choice. In studies of social learning, the size of the selection set is very often limited to just two options. 
For example, in empirical studies of \emph{bombus terresteris} (bumble
bee)~\cite{Baracchi2018} and \emph{parus major} (great tit)~\cite{Aplin2017},
experimenters provided two behaviors for
the bees or birds to choose from, where one yielded a higher payoff than the other. %respectively, each yielding greater or smaller payoffs depending on experimental treatment. 
Many human studies have also used just two
or three possible behaviors~\cite{McElreath2005,Toyokawa2019}.
Modeling studies have often used just two 
behaviors~\cite{Rogers1988,Feldman1996,Rendell2010,perreault2012bayesian}, though others have studied systems with larger selection sets where the number of options is usually not explicitly specified~\cite{boyd1995does,Enquist2007}. It is very rare, however, to find the size of the selection set manipulated as a measure of changing uncertainty.

\textbf{Payoff ambiguity.}
Many models of social learning differentiate between the payoffs for adopting optimal vs. non-optimal behaviors \cite{Rogers1988,Enquist2007,Rendell2010}. The size of the difference between these payoffs is usually taken to indicate the strength of selection of learning, which makes sense. However, in reality payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small payoff other times \cite{McElreath2005}. This means that signals about the relationship between behavior and payoff are often noisy, and differentiating between behavioral options is in part a problem of signal detection.
When the difference in expected payoffs between optimal and non-optimal behaviors is very large, this noise matters little, as the signal is still very clear. However, when the expected payoffs of different behaviors are similar relative to the size of their variances, ambiguity arises about which behaviors really are superior. Smaller differences between payoffs corresponds to larger ambiguity.    
Importantly, payoff ambiguity %differs from selection set size in that it not only affects uncertainty, it also directly
influences both uncertainty \emph{and} the strength of selection. %This parameter sometimes takes the form of differences in expected payoffs arising from different behaviors or strategies~\cite{Enquist2007,Rendell2010} or by varying the standard deviation of payoffs~\cite{McElreath2005}.

\textbf{Effective lifespan.}
Learning can be viewed as a process by which an individual reduces or otherwise manages their uncertainty \cite{jacobs2011bayesian,clark2013whatever}. It is an iterative process by which one repeatedly acquires information, forms representations and predictions, and tests and refines those representations and predictions. It therefore stands to reason that the more opportunities one has for learning, the more one can in principle reduce one's uncertainty. Similarly, a reduction in the number of opportunities to learn will increase the uncertainty about whether one has accurately learned about the available behavioral options and their associated payoffs. We refer to the number of learning opportunities as an individual's effective lifespan to imply that it is the number of opportunities to learn throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities. Empirically, the age of an individual will often correspond with the number of learning opportunities they have had.  Aplin et al. (2017) \nocite{Aplin2017} found that younger great tits more readily used social information compared with older individuals, possibly because they had  accrued less information via individual learning, and also possibly because younger individuals have the most to gain (given their expected remaining lifespan) by switching to superior behavioral options. \ps{I got rid of the Leris and Reader reference because it didn't seem relevant to age-based uncertainty. They just showed that young guppies exposed to reliable models were more likely to use social information later. I also added some human studies, which seems important.} Cross-cultural studies on humans have shown that young children are more likely to acquire their beliefs and simple skills from their parents than are older children or adults, which is also at least partly due to the differential knowledge accrual between young children and the older adults to whom they direct the most trust and attention \cite{kline2013teaching, Reyes2016}. \cm{not sure the second phrase is warranted. Can't seem to find whether social learning overall is  }
%The age of individuals was found to have an effect on social learning in both \emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata} (guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure }
%\mt{Need better connection between lifespan and age, or better references to support use of lifespan}.
Individuals with shorter effective lifespans will end their lives with more uncertainty about which behaviors are optimal, and thereby transmit some of that uncertainty to anyone who learns from them. The effective lifespan for learning varies both across individuals and across species, yet there is little formal theory incorporating this form of uncertainty into our understanding of the evolution of social learning.

@@

\subsection{Varieties of learning, social and otherwise} (CRISTINA) 
\cm{I think a lot of this may be too much detail. I'm still editing and focusing the above on the main forms of uncertainty we are modeling and reviewing the relevant literature as we go. Yes we've made choices about things like horizontal vs vertical, but i think we can just justify that when we introduce the model }
\ps{I agree with Cristina here. I think we should use this section to justify our way of modeling individual and social learning, which is different from previous models. We don't need to provide a detailed review, just why it's worth doing it as we've done it. In particular, we should set up (1) social learning as an initial scaffold for future social learning, and (2) softmax individual learning which involves some greediness with occasional trial and error.}
To add to confusion about how to integrate diverse models of social learning
evolution, different models select different social learning components in
seemingly \emph{ad hoc} ways 
\begin{itemize}
  \item 
    Vertical, oblique, horizontal transmission
  \item
    Conformity, payoff-biased, frequency-dependent, etc. A note about how
    ``conformity'' is often not distinguished from other forms of social learning
  \item
    Review human studies
  \item
    Review non-human studies~\cite{Leris2016,Aplin2017,Avargues-Weber2018,Baracchi2018}
  \item
    One sentence on how dual-inheritance theory enables us to gloss over whether
    social learning is genetically or culturally evolved.
  \item
    More things from which we selected the operation of our model
\end{itemize}

Social learning is the adaptive icing on the cake of intelligent adaptations to ecological uncertainty. To deal with uncertainty as individuals, human experiment participants have been observed to update their tendency for exploring alternative behaviors based on how much uncertainty there is globally. Furthermore, correlations in uncertainty between potential behaviors, humans tend to adopt directed exploration strategies that favor testing behaviors with greater observed payoff variance \cite{Wilson2014,Gershman2019}. Many models of social learning assume little to nothing about the cognitive capabilities of their agents (Cristina unpack this in another paragraph, using the uncertainty studies spreadsheet?). We do not do this because \emph{blackboxing}, or ignoring, cognitive mechanisms is “bad” \cite[p. 658]{Heyes2016, Kendal2018}, but because it is theoretically meaningful. In order to understand how social learning evolves as an adaptation to uncertainty, the model is strengthened by assuming a more biologically plausible, and evolutionarily accurate, model of cognition capable of maximizing payoffs under uncertainty for an individual. In our model we do not impose any structure in randomness between behaviors, so our cognitive model is only sensitive to total randomness. Specifically, we will adopt the softmax distribution for function for selecting a behavior based on its observed payoffs–greater payoff variance means any one behavior is more likely to be selected.

\paragraph{Meaning of uncertainty in ecology (and cognitive and social sciences?) } (JAMIE, PAUL?) \ps{I've added a bunch to the section on different types of uncertainty, which could be more fleshed out if we like. I don't think an additional section here is needed.}
We need to support our claim that these four classes of uncertainty really 
count as forms of uncertainty and show how it is consistent across ecology,
and maybe connect to uncertainty in cognitive and social sciences more broadly.

\paragraph{Evolutionary ABMs for social learning evolution theory development} (PAUL)
\ps{I'm not sure a section is needed on this, but we can add something about how ABMs allow us to capture complexity not possible with purely mathematical approaches. I think it probably makes the most sense to address this at the beginning of the model description.}

@@

\paragraph{Disc. 2: Foundation for studying social learning evolution with group
structure or other topics(PAUL, JAMIE, CRISTINA)}~\cite{Katsnelson2014}.

@@

>>>>

\usepackage{graphicx}
\usepackage{comment}

@@

\section{Introduction}

Social learning enhances problem solving when acquiring information from others is more efficient than learning on one's own~\cite{Laland2004}. However, social learning can also lead individuals astray if, for example, they are copying outdated information. Theoretical models have helped clarify the circumstances under which social learning should be evolutionarily favored ~\cite{BoydRicherson1985}. This theoretical work has motivated much empirical research and helped predict social learning behavior in humans and other species ~\cite{McElreath2005,Kendal2018,Allen2019}.

One proposal arising from this literature maintains that social learning is as an adaptation for dealing with uncertainty, and one that humans heavily rely on. Uncertainty weighs particularly heavily on human adaptation because of our long lifespans, cosmopolitan distribution, and the resulting necessity to adapt to relatively coarse-grain environments \cite{levins1962}. However, the term ``uncertainty" often conflates several different interpretations of that word. For example, different models define uncertainty in terms of the rate of environmental change, spatial environmental variation, or the reliability of information acquired from the environment. Furthermore, in translating mathematical models to verbal predictions, the formal meaning of terms like uncertainty is often lost \cite{lawson1988probability}, facilitating such conflation. As the number of formal models of social learning have expanded, an ever increasing number
of modeling choices ~\cite[Figure 1]{Kendal2018} and formalizations of uncertainty have made it difficult to compare across models and consolidate our understanding of the contexts in which social learning is adaptive. Finally, while risk and uncertainty are often used interchangeably to refer to non-deterministic outcomes, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown \cite{knight1921, volz2012}.

Other (asocial) forms of learning may also be adaptations to uncertainty, but are usually under-emphasized in this literature. Because cultural evolutionary models are more often used to explain social learning mechanisms, they tend to use a single, and often under-powered individual-learning mechanism for comparison. This modelling decision may underestimate the extent to which individual learning can address the adaptive challenges solved by social learning, but also underestimate the quality of information available for social learning.

To address these concerns we developed an evolutionary agent-based model that simultaneously operationalizes several
different forms of uncertainty and endows agents with a powerful learning mechanism. We have chosen to model four kinds of uncertainty that are common in cultural evolution models and the related empirical literature:  (1) temporal environmental variability; (2) selection set size (the number of possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for different behaviors); and (4) the effective lifespan of agents (the number of opportunities for individual learning).\ps{citation needed}. The model also involves a more cognitively realistic and empirically-motivated learning mechanism, the softmax algorithm. This allows a more accurate assessment of tradeoffs between social and individual learning by fleshing out a learning mechanism that makes full use of the individually and socially learned data. We use this integrated model to ask what kinds of uncertainty are likely to favor social learning, and to clarify the logic of why this would be the case.

In the remainder of this introduction we describe previous research on the forms of uncertainty and the learning mechanism we include in our model. We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

%% \cm{from Jamie's intro, not sure we should incorporate the rest here since our model doesn't speak much to things like heavy tailed distributions or adaptive toolkit work: Despite the technical difficulties of dealing with uncertainty, addressing it is an essential task because it potentially changes our understanding of decision-making in a qualitative manner. The dominant approach to understanding decisions for choices with variable outcomes is expected utility theory (EUT). As the name suggests, EUT requires the ability to calculate expectations, which means that we must know the probability distributions of outcomes. The non-probabilistic nature of uncertainty means one cannot calculate expectations. Following \cite{savage1954}, various research traditions in decision theory and economics have treated uncertainty as if it were risk, just with subjective probabilities replacing objective ones. The fundamental problem with this approach was noted by \cite{geweke2001} and further developed by \cite{weitzman2009}. In particular, the probability density function of an outcome that needs to be learned, but for which opportunities for learning are highly constrained, will be heavy-tailed and expected-utility calculations will not work because of the resulting undefined mass-generating function. Uncertainty implies heavy-tailed probability distributions and heavy-tailed distributions require different analytical methods \cite{nair_etal2022}....para...The heuristic-based adaptive-toolbox approach of \cite{todd_gigerenzer2000} provides an alternative strategy for understanding decision-making under uncertainty. There is a strong intuition that uncertainty should activate specific social-learning heuristics. Within the cultural-evolution literature, uncertainty is generally thought to lead to a social heuristic known as \emph{conformist-biased learning} \cite{boyd_richerson1985, henrich_boyd1998, muthukrishna_etal2016}....para...True uncertainty arises when an organism lacks the capacity to learn about the possible outcomes of a decision task. This is very likely for a cosmopolitan species living in a coarse-grain environment. Task complexity also. Non-stationary environmental change \cite{kay_king2020}. These potential sources of uncertainty can also interact.}

@@

\textbf{Temporal environmental variability.}
When external or internal shocks (e.g., climatic events, migration, technological paradigm shifts) occur, strategies that were previously adaptive may no longer be optimal. The difficulty in predicting either when such shocks will occur or what behaviors will be adaptive in the resulting environments leads to uncertainty.

Temporal environmental variability is fundamental to most evolutionary models of social learning. Such fluctuations have been proposed as an important selective pressure for learned as opposed to genetically fixed behaviors ~\cite{Richerson2000}. On the other hand, environments that change too quickly will lead the population to have outdated, and therefore maladaptive information, favoring individual updating over social learning point~\cite{Feldman1996, BoydRicherson1985}. This suggests that intermediate levels of temporal variation are important for the evolution of social learning mechanisms ~\cite{aoki2005}.

Despite the centrality of temporal fluctuations, even this construct is implemented in several ways \cite{aoki2014evolution}. Most commonly, it is modeled as an independent probability of the environment changing its state each generation ~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}. The probabilities of environmental change are usually fixed per generation, %~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
but they have also been modeled as deterministic cycles ~\cite{Feldman1996, aoki2014evolution}. Furthermore, the consequences of environmental change can range from mild to  catastrophic. In the latter case, a change of environment results in the total elimination of any adaptive behavior, which must be learned \emph{de novo} for the new environment ~\cite{Rogers1988}. Such catastrophic environmental changes present a large adaptive challenge as individuals cannot bank on accumulated information from previous generations. Other models of environmental change introduce less uncertainty as they fluctuate between two environmental states with corresponding adaptive behaviors. This means that all individuals with previously maladaptive behaviors then have adaptive behaviors upon the environment changing ~\cite{perreault2012bayesian}. Which strategy one chooses for modelling temporal change can have consequences for at least some of our inferences  ~\cite{aoki2014evolution}.

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options. This is a mathematical fact at the core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the selection set doubles in size  \cite{mackay2003information}. 
This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice.

In studies of social learning, the size of the selection set is very often limited to just two options. 
For example, in empirical studies of \emph{bombus terresteris} (bumble bee)~\cite{Baracchi2018} and \emph{parus major} (great tit)~\cite{Aplin2017}, experimenters provided two behaviors for
the bees or birds to choose from, where one yielded a higher payoff than the other. %respectively, each yielding greater or smaller payoffs depending on experimental treatment. 
Similarly, many human studies have used just two
or three possible behaviors~\cite{McElreath2005,Toyokawa2019}. The simpler empirical choices are often motivated by modeling studies with similar formulations ~\cite{boyd1995does, Rogers1988,perreault2012bayesian}.\cm{pretty sure boyd1995 uses same trick as rogers 88 so not sure why they were being used as illustrating separate kinds of behavior selection sets. correct if wrong} Larger, or more open sets of behavioral choices are not uncommon in experiments trying to capture more complex or naturalistic tasks ~\cite{derex2013, wasielewski2014}, but these map on imperfectly to the theoretical literature. Some models have studied systems with larger, but defined, selection sets ~\cite{Rendell2010}, while in others the number of options is subsumed under probabilities of getting a high payoff ~\cite{Enquist2007}. Researchers have also compared models with a different number of environment states, corresponding to a different number of optimal behaviors through time ~\cite{Feldman1996}, but it is very rare to find the size of the selection set manipulated \emph{within} generations as a measure of changing uncertainty in cultural evolutionary models.

\textbf{Payoff ambiguity.}
Many models of social learning differentiate between the payoffs for adopting optimal vs. non-optimal behaviors \cite{Rogers1988,Enquist2007,Rendell2010}. The size of the difference between these payoffs is usually taken to indicate the strength of selection of learning, which makes sense. However, in reality payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small payoff other times \cite{McElreath2005}. This means that signals about the relationship between behavior and payoff are often noisy, and differentiating between behavioral options is in part a problem of signal detection.
When the difference in expected payoffs between optimal and non-optimal behaviors is very large, this noise matters little, as the signal is still very clear. However, when the expected payoffs of different behaviors are similar relative to the size of their variances, ambiguity arises about which behaviors really are superior. Smaller differences between payoffs corresponds to larger ambiguity. Payoff ambiguity has been manipulated in both theoretical ~\cite{perreault2012bayesian} and empirical ~\cite{McElreath2005l, Morgan2012} studies, both of which support the claim that payoff ambiguity increases the reliance on social information. Importantly, payoff ambiguity %differs from selection set size in that it not only affects uncertainty, it also directly
influences both uncertainty \emph{and} the strength of selection. %This parameter sometimes takes the form of differences in expected payoffs arising from different behaviors or strategies~\cite{Enquist2007,Rendell2010} or by varying the standard deviation of payoffs~\cite{McElreath2005}.

\textbf{Effective lifespan.}
Learning can be viewed as a process by which an individual reduces or otherwise manages their uncertainty \cite{jacobs2011bayesian,clark2013whatever}. It is an iterative process by which one repeatedly acquires information, forms representations and predictions, and tests and refines those representations and predictions. It therefore stands to reason that the more opportunities one has for learning, the more one can in principle reduce one's uncertainty, assuming the environment has not changed in the meantime. Similarly, a reduction in the number of opportunities to learn will increase the uncertainty about whether one has accurately learned about the available behavioral options and their associated payoffs. We refer to the number of learning opportunities as an individual's effective lifespan to imply that it is the number of opportunities to learn throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab, but will also correspond with an individual's age the real world. In multi-round studies of information use in novel tasks, US participants' use of social information declined precipitously across rounds ~\cite{McElreath2005}, suggesting they were more likely to use social information when they were most uncertain about the task early on. In a more naturalistic context, Aplin et al. (2017) \nocite{Aplin2017} found that younger great tits more readily used social information compared with older individuals, possibly because they had accrued less information via individual learning, and also possibly because younger individuals have the most to gain (given their expected remaining lifespan) by switching to superior behavioral options. \ps{I got rid of the Leris and Reader reference because it didn't seem relevant to age-based uncertainty. They just showed that young guppies exposed to reliable models were more likely to use social information later. I also added some human studies, which seems important.} Cross-cultural studies on humans have shown the importance of childhood as a phase of heavy social learning ~\cite{Reyes2016}. Young children are more likely to acquire their beliefs and simple skills from their parents than are older children or adults, which is at least partly due to the differential knowledge accrual between young children and the older adults to whom they direct the most trust and attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both \emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata} (guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure }
%\mt{Need better connection between lifespan and age, or better references to support use of lifespan}.
Individuals with shorter effective lifespans will end their lives with more uncertainty about which behaviors are optimal, and thereby transmit some of that uncertainty to anyone who learns from them. The effective lifespan for learning varies across tasks, individuals, and species, yet most models assume one learning opportunity per generation ~\cite{Feldman1996,Henrich1998, perreault2012bayesian}. Some models do allow several cultural generations within 1 genetic generation ~\cite{Enquist2007}, but little formal theory explicitly examines the role of uncertainty based on number of learning opportunities into our understanding of the evolution of social learning.

@@

\subsection{Other adaptations to uncertainty}

\begin{comment}

\cm{I think a lot of this may be too much detail. I'm still editing and focusing the above on the main forms of uncertainty we are modeling and reviewing the relevant literature as we go. Yes we've made choices about things like horizontal vs vertical, but i think we can just justify that when we introduce the model }
ps{I agree with Cristina here. I think we should use this section to justify our way of modeling individual and social learning, which is different from previous models. We don't need to provide a detailed review, just why it's worth doing it as we've done it. In particular, we should set up (1) social learning as an initial scaffold for future social learning, and (2) softmax individual learning which involves some greediness with occasional trial and error.}

To add to confusion about how to integrate diverse models of social learning
evolution, different models select different social learning components in
seemingly \emph{ad hoc} ways 
\begin{itemize}
  \item 
    Vertical, oblique, horizontal transmission
  \item
    Conformity, payoff-biased, frequency-dependent, etc. A note about how
    ``conformity'' is often not distinguished from other forms of social learning
  \item
    Review human studies
  \item
    Review non-human studies~\cite{Leris2016,Aplin2017,Avargues-Weber2018,Baracchi2018}
  \item
    One sentence on how dual-inheritance theory enables us to gloss over whether
    social learning is genetically or culturally evolved.
  \item
    More things from which we selected the operation of our model
\end{itemize}

\end{comment}

Social learning is the adaptive icing on the cake of other cognitive adaptations to ecological uncertainty. 
While the literature reviewed above suggests that several forms of uncertainty play a role in favoring the evolution of social learning, several other cognitive mechanisms have also evolved in response to uncertainty ~\cite{volz2012}. Many of these are flexible learning mechanisms that do not require assessing others' information. For example, people in experimental studies increase their exploratory behavior when total uncertainty in the task increases and are more likely to test %Furthermore, correlations in uncertainty between potential behaviors, humans tend to adopt directed exploration strategies that favor testing 
behaviors with greater observed payoff variance \cite{Wilson2014,Gershman2019}.

Many models of social learning assume little, or nothing, cognitive capabilities of their agents. For example, a common modelling strategy compare the payoffs of agents with different pure learning strategies (some will always engage in social learning, and others in individual learning) while revealing nothing about the cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988, aoki2005}. Some more complicated learning strategies that make use of both socially and individually learned information have been proposed ~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to reflect cognitive mechanisms for integrating information. Though such simplifications are often useful, several authors have warned of the risks to \emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016, Kendal2018}, given such details may be theoretically meaningful. In order to understand how social learning evolves as an adaptation to uncertainty, we need to endow agents with a  biologically plausible cognition capable of integrating information from various sources in a payoff-maximizing way. \cm{I don't understand this next sentence, but think it can be cut with no loss of meaning: In our model we do not impose any structure in randomness between behaviors, so our cognitive model is only sensitive to total randomness.} To do so, we let agents see the payoffs to behaviors and use the softmax function to convert these observations to a probability distribution from which they strategically select their own behavior. \cm{does this still make sense? and does it need a citation?}

%\paragraph{Meaning of uncertainty in ecology (and cognitive and social sciences?) } (JAMIE, PAUL?) \ps{I've added a bunch to the section on different types of uncertainty, which could be more fleshed out if we like. I don't think an additional section here is needed.}
%We need to support our claim that these four classes of uncertainty really count as forms of uncertainty and show how it is consistent across ecology, and maybe connect to uncertainty in cognitive and social sciences more broadly.

%\paragraph{Evolutionary ABMs for social learning evolution theory development} (PAUL)
%\ps{I'm not sure a section is needed on this, but we can add something about how ABMs allow us to capture complexity not possible with purely mathematical approaches. I think it probably makes the most sense to address this at the beginning of the model description.}

@@

% \paragraph{Disc. 2: Foundation for studying social learning evolution with group
% structure or other topics(PAUL, JAMIE, CRISTINA)}~\cite{Katsnelson2014}.

@@

Taking the functional form of uncertainty seriously is critical for understanding how humans adapt to an uncertain and rapidly changing world, both to mitigate existential threats~\cite{Moya2020,Jones2021}, and to capitalize on new opportunities such as transitioning to clean energy use and production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.
